Metal Mixing and Ejection in Dwarf Galaxies is Dependent on Nucleosynthetic Source

### 2018-07-29 main.tex
section 3.1 revision

\author{Andrew Emerick}
\affiliation{Department of Astronomy, Columbia University, New York, NY, 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY, USA}
\author{Greg L. Bryan}
\affiliation{Department of Astronomy, Columbia University, New York, NY, 10027, USA}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY, 10003, U.S.A}
\author{Mordecai-Mark Mac Low}
\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY, USA}
\author{Benoit C{\^o}t{\'e}}
\affiliation{T.B.D.}
\author{Kathryn V. Johnston}
\affiliation{Department of Astronomy, Columbia University, New York, NY, 10027, USA}
\author{Brian O'Shea}
\affiliation{T.B.D.}

\begin{abstract}
Using a high resolution simulation of an isolated dwarf galaxy, accounting for
%mm
    multi-channel
stellar feedback and chemical evolution on a star-by-star basis, we investigate how each of 15 metal species are distributed within our multi-phase interstellar medium (ISM) and ejected from our galaxy
%mm through
    by
galactic winds. For the first time, we demonstrate that the mass fraction probability distribution functions (PDFs) of individual metal species in the ISM are well described by a piecewise log-normal
%mm +
   and
power-law distribution. The log-normal component generally describes gas at lower metal fractions undergoing enrichment towards the mean metal fraction, driven by recent enrichment events in the power-law tail. The PDF properties, and relative significance of these two components, vary within each ISM phase. Hot gas is dominated by recent enrichment, with a significant power-law tail to high metal fractions, while cold gas is dominated by the log-normal component. In addition, elements dominated by asymptotic giant branch (AGB) wind enrichment (such as N and Ba) mix less efficiently than elements dominated by supernova enrichment (such as $\alpha$ elements and Fe). This difference is driven by the differences in
%mm energetics between each source, particularly the poor coupling of AGB winds to the ISM.
    source locations, particularly the higher chance compared to massive stars for AGB stars to eject material into cold gas.
Nearly all of the produced metals are ejected from the galaxy (only 4\% are retained), but metals dominated by AGB enrichment are retained at a fraction of 20\%. In dwarf galaxies, therefore, elements synthesized predominately through AGB winds should be both overabundant and have a larger spread than elements synthesized in either core collapse or Type Ia supernovae. We discuss the observational implications of these results, their potential use in developing improved models of galactic chemical evolution, and their generalization to more massive galaxies.
\end{abstract}

@@

%mm \textit
\paragraph{Chemistry, Heating, and Cooling Physics} We use the the astrophysical chemistry and cooling package \textsc{Grackle} \citep{GrackleMethod} to evolve a nine species non-equilibrium primordial chemistry model (including molecular hydrogen), follow approximate metal line cooling using a \textsc{Cloudy} look-up table, and apply heating from a metagalactic UV background \citep{HM2012}. We account for approximate self-shielding of H~{\sc i} against the UV background following \cite{Rahmati2013}. We assume He~{\sc i} self-shields in the same fashion as H~{\sc i}, and ignore He~{sc ii} heating from the UVB entirely. Approximate H$_2$ self-shielding from background Lyman-Werner (LW) radiation is accounted for using the Sobolov-like method from \cite{Wolcott-Green2011}. Finally, we use the updated metal line cooling tables which self-consistently account for the decrease in metal line cooling rates due to lower ionization fractions in self-shielding gas, as compared to metal cooling tables computed under the optically thin assumption.

%mm \textit
\paragraph{Star Formation} Stars are followed as individual star particles from 1~M$_{\odot}$ to 100~M$_{\odot}$. Stars are able to form in dense gas with: 1) n~$>$~200~cm$^{-3}$, 2) T~$<$~200~K, and 3) $\nabla \cdot v < 0$. Given the short time-steps ($dt \sim 500$~yr) and high resolution (1.8~pc) in these simulations, the local star formation rate in any single zone is very small ($\ll$ 1~M$_{\odot}$ dt$^{-1}$). We therefore form stars stochastically, depending upon the local gas mass, free-fall time, and star formation efficiency, $\epsilon_{\rm f}$, taken to be 2\%. Stellar masses are randomly sampled from an assumed \cite{Salpeter1955} IMF with metallicities and metal fractions set by the local gas environment. We use the zero age main sequence properties of stars from the \textsc{PARSEC} stellar evolution tables \citep{Bressan2012,Tang2014} to determine individual stellar lifetimes and properties which are used to set their LW band, far ultraviolet (FUV) band, and ionizing radiation luminosities (see below).

%mm \textit
\paragraph{Stellar Feedback and Stellar Yields} We track the feedback and yields of 15 metal species for each star individually. Stars between 8 M$_{\odot} < M_{*} < 25 M_{\odot}$ explode as core collapse supernovae at the end of their life, injecting their mass and 10$^{51}$~erg of thermal energy into a spherical region with radius of 5.4~pc, or 3 times the maximum resolution. Stars above this mass are assumed to direct collapse with no mass or energy injection. For all stars above 8~M$_{\odot}$ we follow their stellar winds assuming continuous mass loss over their lifetimes, their LW and FUV radiation as optically thin radiation which contributes to H$_2$ dissociation and photoelectric heating respectively, and their H~{\sc i} and He~{\sc i} ionizing radiation using an adaptive ray-tracing radiative transfer method \citep{WiseAbel2011}. We interpolate over the OSTAR2002 \citep{Lanz2003} grid to set the luminosities of each of these stars. Low mass stars, $M_{*} < 8~M_{\odot}$, do not produce feedback during their main sequence lifetimes, but end their lives injecting a short-lived, low velocity (10 km~s$^{-1}$) AGB wind. Stars with $3 < M_{*,o} < 8$ are tracked after their death as possible Type Ia supernova progenitors, using a delay time distribution model to assign when (if at all) they will explode as a Type Ia, injecting 10$^{51}$~erg of thermal energy with yields. Stellar yields are computed using the NuGrid stellar yield database \citep{Ritter2017} for all stars with $M_{*} < 25~M_{\odot}$, \textbf{Slemer et. al. in prep} for the stellar winds of stars with $M_{*} > 25~M_{\odot}$, and \cite{Thielemann1986} for Type Ia supernovae.

\paragraph{Initial Conditions} Our dwarf galaxy is initialized to approximate, but not reproduce, the $z = 0$ properties of an ultrafaint dwarf galaxy (UFD) as informed by the observed properties of Leo P \citep[see ][]{Giovanelli2013,McQuinn2015a,McQuinn2015}. We initialize a M$_{\rm gas} = 1.8 \times 10^{6}$~M$_{\odot}$ disk as an expotential profile with $Z = 4.3 \times 10^{-4}$ centered on a static \cite{Burkert1995} dark matter potential with M$_{\rm vir} = 2.5 \times 10^{9}$. The gas scale radius and scale height are set to 250~pc and 100~pc respectively, with a maximum radial extent of 600~pc. Both the gas temperatures and velocities are set iteratively to enforce initial hydrostatic equilibrium. The galaxy contains no initial background stellar population, limiting the number of Typa Ia supernovae in our model, but we do provide initial supernova driving to limit the initial transient burst of star formation from the initial cooling and collapse of the galaxy. Although the total metallicity field is initially non-zero, we only track the self-consistently produced metal enrichment for each individual metal species in our simulation, setting their initial abundances to zero. Our analysis is based on the evolution of this galaxy during the first 500~Myr after the formation of the first star particle.

\section{Results}
For context, in Paper I we focused on the global properties of the evolution of this dwarf galaxy, including an analysis of the galaxy's gas mass and star formation evolution, the ISM properties in terms of mass fractions, volume fractions, and phase diagrams, the interstellar radiation field in each tracked radiation band, the gas outflow rates and galactic wind velocities, and the retention / ejection of metals from the galaxy. This galaxy has an average star formation rate of $1.2~\times 10^{4}$ M$_{\odot}$ yr$^{-1}$ and is consistent with observed low mass dwarfs galaxies in the Kenicutt-Schmidt relation.  The galaxy exhibits strong, feedback driven outflows that eject a significant amount of gas and metals from the galaxy. The mass loading factor at 0.25~R$_{\rm vir}$ was found to be $\eta \sim 50$. These winds eject 95\% of the metals produced in the galaxy, with 50\% of all metals leaving the virial radius by the end of the simulation time.

@@

The top panel of Figure~\ref{fig:species_fractions} gives the mass fraction of each element at the end of the simulation (500~Myr) in each of four reservoirs: locked in stars (yellow), in the ISM (blue), outside the galaxy but within the virial radius (purple), and outside the virial radius (salmon), including gas that has left the domain. We subdivide the ISM by phase in the bottom panel, giving the mass fractions of each element in the cold neutral medium (CNM, $f_{\rm H_2} < 0.5$,  T$< 100$~K), warm neutral medium (WNM, $10^2~\rm{K}\le \rm{T} < 10^4~\rm{K}$), warm ionized medium (WIM, $10^4~\rm{K}\le \rm{T} < 10^{5.5}~\rm{K}$), the hot ionized medium (H~{\sc i}M T$\ge 10^{5.5}$~K), and locked in stars (yellow). Again, with the exception of H and He, we only consider metals produced self-consistently through our star formation and stellar feedback methods; the initial mass of each metal species is zero.

We find two major results. First, nearly all of the metals reside within neutral gas, mostly in the CNM; only a few percent reside in the hot phases. The exact fraction varies with each species, most notably for carbon; these fluctuations are at most $\sim 10$\%. This is not surprising, as the cold phases represent the majority of the mass in the ISM, but, as shown in Section~\ref{sec:statistics}, even though the cold phases harbor most of the metals, the hot phases have significantly higher metal mass fractions. Second, only a small fraction of produced metals are retained within the dwarf galaxy, in agreement with observations of nearby dwarf galaxies \citep[see][]{Kirby2011-metals, McQuinn2015}; however,the retention factor varies. The top panel shows a qualitative disagreement between the retention fractions of N and Ba (about 20\% for each) as compared to the rest of the metals ($\sim$ 4 -- 5\%). This suggests that individual metals \textit{do not} share the same dynamical evolution. Clearly, metal enrichment in galaxies is a phenomenon that cannot be fully captured using single, global metallicity field. In this particular case, assuming that all metals behave the same would underestimate the N and Ba enrichment by a factor of up to five, at least for low mass dwarf galaxies.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{species_bar_sources.png}\\
\caption{The fraction of total mass in each metal species produced by each of the four possible nucleosynthetic channels in our model. These channels differ in both when metals are ejected, as determined by stellar evolution, and the energy with which they are ejected into the ISM. We note that the minimal contribution from Type Ia supernovae for the iron-peak elements is because only XX of them have exploded by the end of our 500~Myr simulation.}
% GB(AE): Why are H and He different? Either discuss or remove from plot.
%  ------ AE removed from plot
%
\label{fig:species_sources}
\end{figure}

The only physics that separates the dynamical evolution of these elements in our simulations is the individual sources of enrichment: AGB winds (stars less than 8~M$_{\odot}$), stellar winds (stars above 8~M$_{\odot}$), core collapse supernovae, and Type Ia supernovae. These channels differ in: 1) how long after a given star formation event they occur, 2), by consequence, the typical ISM properties in which they occur, 3) how much energy is associated with each event, which determines ejecta temperature and velocity. To understand where each of the elements in Figure~\ref{fig:species_fractions} originated, we show the mass fraction of metals produced through each channel in Figure~\ref{fig:species_sources}. Core collapse supernovae are responsible for over 90\% of the total metal enrichment in our galaxy, but this is clearly not true for all elements. In particular, a majority of N (74\%) and Ba (65\%) are ejected by AGB winds; both are retained at a higher rate than the rest of the metals. That this behavior exists for N and Ba in our simulations is dependent upon the metallicity of our galaxy and choice of stellar yields tables, but we can generalize this result to say that, for any assumed set of yields, low mass galaxies should more easily retain \textit{any} elements synthesized predominately in AGB winds, as compared to elements synthesized through supernovae.\footnote{One might expect that much of the N and Ba that is ejected by galactic winds is comprised mostly of the $\sim$30\% of each species produced through the stellar winds of more massive stars and supernovae. However, this cannot be determined as we lack Lagrangian information about individual gas elements.}
% AE: Actually if there are significant isotope variations between nucleosynthetic sources (i.e. N produced in AGB winds is a different isotope than the trace N produced in SNe) then that could be a very cool thing to distinguish if we had good gas-phase abundance measures of various isotopes... but maybe not if they aren't very stable....
 We discuss how this result may extend towards other metal yields that are not well sampled on our relatively short (500 Myr) simulation timescales in Section~\ref{sec:discussion:metal yields}.

AGB winds have low energy and velocities (10~km~s$^{-1}$) as compared to the energy and typical expansion velocities of supernovae ($\sim$1000~km~s$^{-1}$). In addition, their longer timescales (40-100~Myr as compared to $\sim$10~Myr) means that AGB stars are typically removed from their birth regions. The changes in typical ISM density and height of these events contributes to these variations. We show histograms of the average number density within 20~pc of any given enrichment source (top) and height above/below the disk (bottom) within 1~Myr of each event (as limited by our output cadence) in Figure~\ref{fig:spatial distribution}. As shown, SN peak at very low densities, indicating that most explode in superbubbles, regions carved out by previous SNe. AGB stars predominately release their metals close to the average ISM density. The scale height distributions for both events show no significant differences. We do not expect these differences to be the dominant effect in driving the differential evolution of elements ejected by AGB winds vs. supernovae, compared to the energetics, but can certainly play a significant role in determining the mixing behavior of individual enrichment events. The changes to mixing behavior as a function of ISM properties will be investigated in more detail in a future work.

@@

Chemical evolution models make varying approximations as to what fraction and on what timescales newly formed metals populate a galaxy's ISM and become available to enrich future generations of stars. In the simplest models, new metals are instantaneously available upon the formation of new stars, ignoring delays due to stellar lifetimes (\textit{refs, but maybe not if included in introduction}). Accounting for delays due to individual stellar lifetimes, loss from galactic winds, and dilution from accretion of pristine gas leads to models that can more accurately capture a variety of mean chemical properties of galaxies with some ability to account for spreads in abundances. In hydrodynamics simulations, there has been substantial work understanding how metals, often characterized as passive scalar fields, diffuse in both turbulent boxes and global galaxy models. However, there is not  yet a complete model for how metals should mix and evolve in the complex, multi-phase ISM of real galaxies. To build towards this, we characterize the metal evolution and distributions in our galaxy below.

@@

%

@@

\subsection{Do these results apply to more massive galaxies?}
\label{sec:massive galaxies}
One important caveat about our work is that these results are derived from simulations of an isolated, low mass dwarf galaxy whose properties vary dramatically from a more massive galaxy like the Milky Way. It is unclear how much of these results apply to more massive galaxies given their deeper potential wells and higher star formation rates. Feedback driven galactic outflow properties do vary significantly with halo mass \citep[e.g.][]{MacLowFerrara1999,Muratov2017}, as more massive galaxies more easily retain and re-accrete gas. We expect the results presented in Figure~\ref{fig:species_fractions} to be the most susceptible to the particular star formation history and halo depth of a given galaxy. With gas from individual enrichment events more easily contained, we would expect the retention fractions to be more similar across metal species with increasing halo mass or decreasing star formation rate. If true, this difference in metal retention can be a key observable in verifying our result that the dynamical evolution of metals in the ISM is not uniform. If massive galaxies retain metals equally, we would expect dwarf galaxies to exhibit larger abundance ratios between AGB wind elements and SN enriched elements than stars in more massive galaxies, like the Milky Way. This difference would be greater at later times (higher metallicity) once AGB enrichment becomes significant. For example, at metallicities of [Fe/H] $\gtrsim$ 1, Ba is predominantly produced via s-process and ejected in AGB winds \citep{Travaglio1999,Travaglio2004}. At this metallicity, we would expect an elevated [Ba/$\alpha$] at fixed [$\alpha$/H] for low mass dwarfs as compared to the Milky Way. Indeed, this is found in the dSph's of the Milky Way \citep[see ][]{Tolstoy2009} and has been argued to either be the result of variations in s-process enrichment in metal poor environments and/or the result of enrichment timescale differences given the different star formation histories of dwarf galaxies and the Milky Way. In addtion, recent observations find elevationd [Ba/Fe] vs. [Fe/H] in local dwarf galaxies \citep{DugganKirbyAAS},
% AE TODO: Update duggand and kirby if / when the paper comes out
which could possibly be explained by this effect, but could also point towards significant Ba enrichment from NSMs. Given the lack of definitive explanation, clearly generalizing our results and detailing their observational consequences is a valuable area of future research.

>>>>

\author{Andrew Emerick}
\affiliation{Department of Astronomy, Columbia University, New York, NY, 10027, USA}
\author{{\it more...}}

\begin{abstract}
Using a high resolution simulation of an isolated dwarf galaxy, accounting for 
%mm 
    multi-channel 
stellar feedback and chemical evolution on a star-by-star basis, we investigate how each of 15 metal species are distributed within our multi-phase interstellar medium (ISM) and ejected from our galaxy 
%mm through 
    by
galactic winds. For the first time, we demonstrate that the mass fraction probability distribution functions (PDFs) of individual metal species in the ISM are well described by a piecewise log-normal 
%mm + 
   and
power-law distribution. The log-normal component generally describes gas at lower metal fractions undergoing enrichment towards the mean metal fraction, driven by recent enrichment events in the power-law tail. The PDF properties, and relative significance of these two components, vary within each ISM phase. Hot gas is dominated by recent enrichment, with a significant power-law tail to high metal fractions, while cold gas is dominated by the log-normal component. In addition, elements dominated by asymptotic giant branch (AGB) wind enrichment (such as N and Ba) mix less efficiently than elements dominated by supernova enrichment (such as $\alpha$ elements and Fe). This difference is driven by the differences in 
%mm energetics between each source, particularly the poor coupling of AGB winds to the ISM. 
     temperature between fast, shocked supernova ejecta and slow AGB winds.
%mm [I agree with the argument below that this is secondary]    source locations, particularly the higher chance compared to massive stars for AGB stars to eject material into cold gas.
Nearly all of the produced metals are ejected from the galaxy (only 5\% are retained), but metals dominated by AGB enrichment are retained at a fraction of 20\%. In dwarf galaxies, therefore, elements synthesized predominately through AGB winds should be both overabundant and have a larger spread than elements synthesized in either core collapse or Type Ia supernovae. We discuss the observational implications of these results, their potential use in developing improved models of galactic chemical evolution, and their generalization to more massive galaxies.
\end{abstract}

@@

%mm \textit
\paragraph{Chemistry, Heating, and Cooling Physics} We use the the astrophysical chemistry and cooling package \textsc{Grackle} \citep{GrackleMethod} to evolve a nine species non-equilibrium primordial chemistry model (including molecular hydrogen), follow approximate metal line cooling using a \textsc{Cloudy} look-up table, and apply heating from a metagalactic UV background \citep{HM2012}. We account for approximate self-shielding of H~{\sc i} against the UV background following \cite{Rahmati2013}. We assume He~{\sc i} self-shields in the same fashion as H~{\sc i}, and ignore He~{\sc ii} heating from the UVB entirely. Approximate H$_2$ self-shielding from background Lyman-Werner (LW) radiation is accounted for using the Sobolev-like method from \cite{Wolcott-Green2011}. Finally, we use updated metal line cooling tables that self-consistently account for the decrease in metal line cooling rates due to lower ionization fractions in self-shielding gas, as compared to metal line cooling tables computed under the optically thin assumption.

%mm \textit
\paragraph{Star Formation} Stars are followed as individual star particles from 1~M$_{\odot}$ to 100~M$_{\odot}$. Stars are able to form in dense gas with: 1) n~$>$~200~cm$^{-3}$, 2) T~$<$~200~K, and 3) $\nabla \cdot v < 0$. Given the short time-steps ($dt \sim 500$~yr) and high resolution (1.8~pc) in these simulations, the local star formation rate in any single zone is very small ($\ll$ 1~M$_{\odot}$ dt$^{-1}$). We therefore form stars stochastically, depending upon the local gas mass, free-fall time, and star formation efficiency, $\epsilon_{\rm f}$, taken to be 2\%. Stellar masses are randomly sampled from an assumed \cite{Salpeter1955} IMF with metallicities and metal fractions set by the local gas environment. We use the zero-age main sequence properties of stars from the \textsc{PARSEC} stellar evolution tables \citep{Bressan2012,Tang2014} to determine individual stellar lifetimes and properties which are used to set their LW band, far ultraviolet (FUV) band, and ionizing radiation luminosities (see below).

%mm \textit
\paragraph{Stellar Feedback and Stellar Yields} We track the feedback and yields of 15 metal species for each star individually. Stars with mass 8~M$_{\odot} < M_{*} < 25$~M$_{\odot}$ explode as core collapse supernovae at the end of their lives, injecting their mass and 10$^{51}$~erg of thermal energy into a spherical region with radius of 5.4~pc, or 3 times the maximum resolution. Stars above this mass are assumed to
%mm direct collapse 
    directly collapse to black holes with no mass or energy 
%mm injection. 
    ejection.
For all stars above 8~M$_{\odot}$ we follow: their stellar winds assuming continuous mass loss over their lifetimes
%mm [this is a rather important approximation that should be explicitly noted]
    but neglecting the energy injection of the winds by reducing their velocities (to 10 
    km~s$^{-1}$); 
their LW and FUV radiation as optically thin radiation which contributes to H$_2$ dissociation and photoelectric heating respectively; and their H~{\sc i} and He~{\sc i} ionizing radiation using an adaptive ray-tracing radiative transfer method \citep{WiseAbel2011}. We interpolate over the OSTAR2002 \citep{Lanz2003} grid to set the luminosities of each of these stars. Low mass stars with $M_{*} < 8$~M$_{\odot}$, do not produce feedback during their main sequence lifetimes, but end their lives injecting a short-lived, low velocity (10 km~s$^{-1}$) AGB wind. Stars with $3 < M_{*,o} < 8$ are tracked after their death as possible Type Ia supernova progenitors, using a delay time distribution model to assign when (if at all) they will explode, injecting 10$^{51}$~erg of thermal energy with yields. Stellar yields are computed using the NuGrid stellar yield database \citep{Ritter2017} for all stars with $M_{*} < 25$~M$_{\odot}$, \textbf{Slemer et. al. in prep} for the stellar winds of stars with $M_{*} > 25$~M$_{\odot}$, and \cite{Thielemann1986} for Type Ia supernovae.

\paragraph{Initial Conditions} Our dwarf galaxy is initialized to approximate, but not reproduce, the 
%mm redshift $z = 0$ 
properties of 
%mm an 
    a modern
ultrafaint dwarf galaxy, as informed by the observed properties of Leo P \citep[see ][]{Giovanelli2013,McQuinn2015,McQuinn2015a}. We initialize a disk with $M_{\rm gas} = 1.8 \times 10^{6} \, M_{\odot}$ distributed in an exponential profile centered on a static \cite{Burkert1995} dark matter potential with 
%mm [added unit of M_odot to M_vir]
M$_{\rm vir} = 2.5 \times 10^{9}$~M$_{\odot}$. 
The gas scale radius and scale height are set to 250~pc and 100~pc respectively, with a maximum radial extent of 600~pc. Both the gas temperatures and velocities are set iteratively to enforce initial hydrostatic equilibrium. The galaxy contains no initial background stellar population, limiting the number of Type Ia supernovae in our model, but we do provide initial 
%mm
   randomly placed 
supernova driving to limit the initial transient burst of star formation from the initial cooling and collapse of the galaxy. Although we set the total metallicity field is initially non-zero, with $Z = 4.3 \times 10^{-4}$, we only track the self-consistently produced metal enrichment for each individual metal species in our simulation, setting their initial abundances to zero. Our analysis is based on the evolution of this galaxy during the first 500~Myr after the formation of the first star particle.

\section{Results}
%mm For context, in 
     In
Paper I we focused on the global properties of the evolution of this dwarf galaxy, including an analysis of the galaxy's gas mass and star formation evolution, the ISM properties in terms of mass fractions, volume fractions, and phase diagrams, the interstellar radiation field in each tracked radiation band, the gas outflow rates and galactic wind velocities, and the
%mm 
   overall rate of 
retention or ejection of metals from the galaxy. This galaxy has an average star formation rate of $1.2~\times 10^{4}$ M$_{\odot}$ yr$^{-1}$ and is consistent with observed low mass dwarfs galaxies in the Kennicutt-Schmidt relation 
%mm
     \citep{Kennicutt1998}.  
The galaxy exhibits strong, feedback-driven outflows that eject a significant amount of gas and metals from the galaxy. The mass loading factor at 0.25~R$_{\rm vir}$ was found to be $\eta \sim 50$. These winds eject 95\% of the metals produced in the galaxy, with 50\% of all metals leaving the virial radius by the end of the simulation time.

@@

The top panel of Figure~\ref{fig:species_fractions} gives the mass fraction of each element at the end of the simulation (500~Myr) in each of four reservoirs: locked in stars (yellow), in the ISM (blue), outside the galaxy but within the virial radius (purple), and outside the virial radius (salmon), including gas that has left the domain. We subdivide the ISM by phase in the bottom panel, giving the mass fractions of each element in the cold neutral medium (CNM, $f_{\rm H_2} < 0.5$,  T$< 100$~K), warm neutral medium (WNM, $10^2~\rm{K}\le \rm{T} < 10^4~\rm{K}$), warm ionized medium (WIM, $10^4~\rm{K}\le \rm{T} < 10^{5.5}~\rm{K}$), the hot ionized medium (H~{\sc i}M T$\ge 10^{5.5}$~K), and locked in stars (yellow). 
%mm Again, with the exception of  H and He, we [H & He aren't metals]
    We
only consider metals produced self-consistently through our star formation and stellar feedback methods; the initial mass of each metal species is zero.

We find two major results.
%mm [reversed order of results to follow figure.  The opposite choice could also be made, by reversing the order of the figure...]
    First, 
only a small fraction of produced metals are retained within the dwarf galaxy, in agreement with observations of nearby dwarf galaxies \citep[see][]{Kirby2011-metals, McQuinn2015}; however, the retention factor varies 
%mm 
    among elements. The top panel shows a qualitative disagreement between the retention fractions of N and Ba (about 20\% for each) as compared to the rest of the metals ($\sim$ 4 -- 5\%). This suggests that individual metals \textit{do not} share the same dynamical evolution. 
%mm Clearly
       Thus, 
metal enrichment in galaxies is a phenomenon that cannot be fully captured using a single, global metallicity field. In this particular case, assuming that all metals behave the same would underestimate the N and Ba enrichment by a factor of up to five, at least for low mass dwarf galaxies.

    Second,
nearly all of the metals 
%mm
    retained in the disk
reside within neutral gas, mostly in the CNM; only a few percent
reside in the hot phases. The exact fraction varies with each species, most notably for carbon; these fluctuations are at most $\sim 10$\%. This is not surprising, as the cold phases represent the majority of the mass in the ISM, but, as shown in Section~\ref{sec:statistics}, even though the cold phases harbor most of the metals, the hot phases have significantly higher metal mass fractions.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{species_bar_sources.png}\\
\caption{The fraction of total mass in each metal species produced by each of the four possible nucleosynthetic channels in our model. These channels differ in both when metals are ejected, as determined by stellar evolution, and the 
%mm energy with which they are ejected into the ISM. 
     phase of the ISM into which they are ejected.
We note that the minimal contribution from Type Ia supernovae for the iron-peak elements is because 
%mm
    no older stellar population was initialized, so 
only 
%mm [boldfaced]
     {\bf XX} 
of them have exploded by the end of our 500~Myr simulation.}
% GB(AE): Why are H and He different? Either discuss or remove from plot.
%  ------ AE removed from plot
%
\label{fig:species_sources}
\end{figure}

The only physics that separates the dynamical evolution of these elements in our simulations is the individual sources of enrichment: AGB winds (stars less than 8~M$_{\odot}$), stellar winds (stars above 8~M$_{\odot}$), core collapse supernovae, and Type Ia supernovae. These channels differ in: 1) how long after a given star formation event they occur, 2), by consequence, the typical ISM properties in which they occur, and 3) how much energy is associated with each event, which determines ejecta temperature and velocity. To understand where each of the elements in Figure~\ref{fig:species_fractions} originated, we show the mass fraction of metals produced through each channel in Figure~\ref{fig:species_sources}. Core collapse supernovae are responsible for over 90\% of the total metal enrichment in our galaxy, but this is clearly not true for all elements. In particular, a majority of N (74\%) and Ba (65\%) are ejected by AGB winds; both are retained at a higher rate than the rest of the metals. That this behavior exists for N and Ba in our simulations is dependent upon the metallicity of our galaxy and choice of stellar yield tables, but we can generalize this result to say that, for any assumed set of yields, low mass galaxies should more easily retain \textit{any} elements synthesized predominately in AGB winds, as compared to elements synthesized through supernovae.\footnote{One might expect that much of the N and Ba that is ejected by galactic winds is comprised mostly of the $\sim$30\% of each species produced through the stellar winds of more massive stars and supernovae. However, this cannot be 
%mm determined 
    proven
as we lack Lagrangian information about individual gas elements.}
% AE: Actually if there are significant isotope variations between nucleosynthetic sources (i.e. N produced in AGB winds is a different isotope than the trace N produced in SNe) then that could be a very cool thing to distinguish if we had good gas-phase abundance measures of various isotopes... but maybe not if they aren't very stable....
 We discuss how this result may extend towards other metal yields that are not well sampled on our relatively short (500 Myr) simulation timescales in Section~\ref{sec:discussion:metal yields}.

AGB winds have low energy and velocities (10~km~s$^{-1}$) as compared to the energy and typical expansion velocities of supernovae ($\sim$1000~km~s$^{-1}$). In addition, their longer timescales (40--100~Myr as compared to $\sim$10~Myr) means that AGB stars are typically removed from their birth regions 
%mm
    and randomly distributed through the galactic disk. 
The changes in typical ISM density and height of these events contributes to these variations. We show histograms of the average number density within 20~pc of any given enrichment source (top) and height above/below the disk (bottom) within 1~Myr of each event (as limited by our output cadence) in Figure~\ref{fig:spatial distribution}. As shown, SN peak at very low densities, indicating that most explode in superbubbles, regions carved out by previous SNe. AGB stars predominately release their metals close to the average ISM density. The scale height distributions for both events show no significant differences. We do not expect these differences to be the dominant effect in driving the differential evolution of elements ejected by AGB winds vs. supernovae, compared to the energetics, but they can certainly play a significant role in determining the mixing behavior of individual enrichment events. The changes to mixing behavior as a function of ISM properties will be investigated in more detail in a future work.

@@

%mm should most of this material be in the introduction, and just the last two sentences be here?
Chemical evolution models make varying approximations as to what fraction and on what timescales newly formed metals populate a galaxy's ISM and become available to enrich future generations of stars. In the simplest models, new metals are instantaneously available upon the formation of new stars, ignoring delays due to stellar lifetimes (\textit{refs, but maybe not if included in introduction}). Accounting for delays due to individual stellar lifetimes, loss from galactic winds, and dilution from accretion of pristine gas leads to models that can more accurately capture a variety of mean chemical properties of galaxies with some ability to account for spreads in abundances. In hydrodynamics simulations, there has been substantial work understanding how metals, often characterized as passive scalar fields, diffuse in both turbulent boxes and global galaxy models. However, there is not  yet a complete model for how metals should mix and evolve in the complex, multi-phase ISM of real galaxies. To build towards this, we characterize the metal evolution and distributions in our galaxy below.

@@

The first core collapse supernova occurs at 12~Myr,

@@

\subsection{Do these results apply to more massive galaxies?}
\label{sec:massive galaxies}
One important caveat about our work is that these results are derived from simulations of an isolated, low mass dwarf galaxy whose properties vary dramatically from a more massive galaxy like the Milky Way. It is unclear how much of these results apply to more massive galaxies given their deeper potential wells and higher star formation rates. Feedback driven galactic outflow properties do vary significantly with halo mass \citep[e.g.][]{MacLowFerrara1999,Muratov2017}, as more massive galaxies more easily retain and re-accrete gas. We expect the results presented in Figure~\ref{fig:species_fractions} to be the most susceptible to the particular star formation history and halo depth of a given galaxy. With gas from individual enrichment events more easily contained, we would expect the retention fractions to be more similar across metal species with increasing halo mass or decreasing star formation rate. If true, this difference in metal retention can be a key observable in verifying our result that the dynamical evolution of metals in the ISM is not uniform. If massive galaxies retain metals equally, we would expect dwarf galaxies to exhibit larger abundance ratios between AGB wind elements and SN enriched elements than stars in more massive galaxies, like the Milky Way. This difference would be greater at later times (higher metallicity) once AGB enrichment becomes significant. For example, at metallicities of [Fe/H] $\gtrsim$ 1, Ba is predominantly produced via s-process and ejected in AGB winds \citep{Travaglio1999,Travaglio2004}. At this metallicity, we would expect an elevated [Ba/$\alpha$] at fixed [$\alpha$/H] for low mass dwarfs as compared to the Milky Way. Indeed, this is found in the dSph's of the Milky Way \citep[see ][]{Tolstoy2009} and has been argued to either be the result of variations in s-process enrichment in metal poor environments and/or the result of enrichment timescale differences given the different star formation histories of dwarf galaxies and the Milky Way. However, there has been no definitive explanation for these observations. Generalizing our results and detailing their observational consequences is the topic of ongoing work.
